Appendix 3: Estimating species interactions

### 2015-11-01 Harris_2015/manuscript-materials/draft.Rmd
and another thing

### Introduction

@@

To the extent that nontrophic species interactions (such as competition) affect
community assembly, ecologists might expect to find signatures of these
interactions in species composition data [@macarthur_population_1958;
@diamond_island_1975]. Despite decades of work and several major controversies,
however [@lewin_santa_1983; @strong_ecological_1984; @connor_checkered_2013],
existing methods for detecting competition's effects on community structure are
unreliable [@gotelli_empirical_2009]. In particular, species' effects on one
another can become lost in a web of indirect effects. For example, the
competitive interaction between the two shrub species in Figure 1A is
obscured by their shared tendency to occur in unshaded areas (Figure 1B). While
ecologists have long known that indirect effects can overwhelm direct ones at
the landscape level [@dodson_complementary_1970; @levine_competitive_1976], the
vast majority of our methods for drawing inferences from observational data do
not control for these effects [e.g. @diamond_island_1975;
@strong_ecological_1984; @gotelli_empirical_2009; @veech_probabilistic_2013;
@pollock_understanding_2014]. To the extent that indirect interactions like
those in Figure 1 are generally important [@dodson_complementary_1970], existing
methods will not provide much evidence regarding species interactions.

@@

### Methods

**Markov networks.**
Markov networks provide a framework for translating back and forth between the
conditional (all-else-equal) relationships among species (Figure 1C) and the
kinds of species assemblages that these relationships produce. Here, I show how
a set of conditional relationships can be used to determine how groups of species
can co-occur. Methods for estimating conditional relationships from data are
discussed in the next section.

@@

### Results

@@

\noindent
\textbf{Twenty species.}
Despite some variability across contexts (Figure 3), the four methods that
controlled for indirect effects (the Markov network, the generalized linear
models, and the two partial correlation-based methods) outperformed those
that did not.  Of these, the Markov network consistently performed the best,
especially on the smaller landscapes [which are also the most
representative of typical analyses in this field; @gotelli_empirical_2009].
As shown in Appendix [[X]], the standard errors associated with the estimates
in Figure 3 are small (less than 0.01), so these differences should not be
attributed to sampling error.

Of the methods that did not control for indirect effects, Figure 3 shows that
simple correlation coefficients provided a more reliable indicator of species'
true interaction strengths than either the joint species distribution model
(BayesComm) or the null model (Pairs). The estimates from these
three approaches were tightly correlated (Figure [[3B]]), suggesting that the
null model only contains a noisy version of the same information that could be
obtained more easily and interpretably with simple correlation coefficients.

Finally, we can evaluate the models' statistical inferences.  Figure 4 shows,
across a range of true interaction strengths, the probability that the null model
or the Markov network will predict the wrong sign of the interaction with 95%
confidence.  The null model makes such errors nearly 8 times as often as the
Markov network, even though it only reject the null hypothesis twice as often
overall [[Appendix]].  The Markov network's errors were also more concentrated
around 0, so it never misclassified strong interactions like the null model did.

### Discussion

The results presented above show that Markov networks can reliably recover
species' pairwise interactions from observational data, even for cases where
environmental heterogeneity and indirect interactions cause ecologists' typical
null modeling approaches to reliably fail. Partial covariances and generalized
linear models can both provide computationally efficient approximations, but
with somewhat lower accuracy. The difference in accuracy may be larger for real
data sets than for the simulated landscapes in Figure 3, however; linear
approximations to the Markov network make larger errors when the interaction
matrix is structured [e.g. due to guilds or trophic levels;
@loh_structure_2013]. Similarly, the separate generalized linear models
for each species can severely overfit in some cases [@lee_learning_2012]. The
full Markov network should thus be preferred to the approximations when
it is computationally feasible.

Compositional data only has enough degrees of freedom to estimate one interaction
per species pair [@schmidt_modeling_2012], so none of these methods can identify
the exact nature of the pairwise interactions [e.g. which species in a
positively-associated pair is facilitating the other]. To estimate asymmetric
interactions, such as commensalism or predation, ecologists could use time
series, behavioral observations, manipulative experiments, or natural history.
These other sources of information could also be used to augment the likelihood
function with an informative prior distribution, reducing ecologists' eror and
uncertainty.

@@

Finally, the results presented here have important implications for
ecologists' continued use of null models for studying species interactions. Null
and neutral models can be useful for clarifying our thinking
[@harris_occupancy_2011; @xiao_strong_2015], but deviations from a particular
null model must be interpreted with care [@roughgarden_competition_1983]. Even
in small networks with three species, it may simply not be possible to implicate
specific ecological processes like competition by rejecting a general-purpose
null [@gotelli_empirical_2009], especially when the test statistic is
effectively just a correlation coefficient [[Figure 3B]]. When a non-null
backdrop is not controlled for, Type I error rates can skyrocket, the apparent
sign of the interaction can change, and null models can lead to reliably
incorrect inferences (Figure 1, Figure 4).

@@

\noindent \textbf{Figure 4: A.} Inferences from the Markov network's 95%
confidence intervals on the "no-environment" simulated landscapes, as a function
of the "true" coefficient values. **B.** Inferences from the $Z$-statistics
estimated by the "Pairs" software for null modeling.

**References:**

>>>>

\noindent\textbf{Introduction}

@@

\noindent
To the extent that nontrophic species interactions (such as competition) affect
community assembly, ecologists might expect to find signatures of these
interactions in species composition data [@macarthur_population_1958;
@diamond_island_1975]. Despite decades of work and several major controversies,
however [@lewin_santa_1983; @strong_ecological_1984; @connor_checkered_2013],
existing methods for detecting competition's effects on community structure are
unreliable [@gotelli_empirical_2009]. In particular, species' effects on one
another can become lost in a web of indirect effects. For example, the
competitive interaction between the two shrub species in Figure 1A is
obscured by their shared tendency to occur in unshaded areas (Figure 1B). While
ecologists have long known that indirect effects can overwhelm direct ones at
the landscape level [@dodson_complementary_1970; @levine_competitive_1976], the
vast majority of our methods for drawing inferences from observational data do
not control for these effects [e.g. @diamond_island_1975;
@strong_ecological_1984; @gotelli_empirical_2009; @veech_probabilistic_2013;
@pollock_understanding_2014]. To the extent that indirect interactions like
those in Figure 1 are generally important [@dodson_complementary_1970], existing
methods will not provide much evidence regarding species interactions.

@@

\noindent\textbf{Methods}

\noindent
\textbf{Markov networks.}
Markov networks provide a framework for translating back and forth between the
conditional (all-else-equal) relationships among species (Figure 1C) and the
kinds of species assemblages that these relationships produce. Here, I show how
a set of conditional relationships can be used to determine how groups of species
can co-occur. Methods for estimating conditional relationships from data are
discussed in the next section.

@@

\noindent
\textbf{Results}

@@

\noindent
\textbf{Twenty species.}
Despite some variability across contexts (Figure 3), the four methods that
controlled for indirect effects (the Markov network, the generalized linear
models, and the two partial correlation-based methods) outperformed those
that did not.  Of these, the Markov network consistently performed the best.
As anticipated by @lee_learning_2012, the Markov network produced very
similar estimates to the generalized linear models (Figure 4A), but with
higher accuracy---especially with smaller data sets.  As reviewed in
@gotelli_empirical_2009, most analyses in this field in ecology involve
fewer than 50 sites, so this is the most important regime for model
performance. As shown in Appendix [[X]], the standard errors associated
with the estimates in Figure 3 are small (less than 0.01), so these
differences should not be attributed to sampling error.

Of the methods that did not control for indirect effects, Figure 3 shows that
simple correlation coefficients provided a more reliable indicator of species'
true interaction strengths than either the joint species distribution model
(BayesComm) or the null model (Pairs). The estimates from these three approaches
were tightly correlated, after controlling for the size of the landscape,
suggesting that the null model only contains a noisy version of the
same information that could be obtained more easily and interpretably with
simple correlation coefficients (Figure 4B).

Finally, we can evaluate the models' statistical inferences (focusing on the
first two simulation types, for which the true interaction rates are easiest to
interpret).  The Markov network's Type I error rate (defined here as its
probability of rejecting the null hypothesis for a pair of species where
$|\beta_{ij}|<0.1$ depended on the simulation type: 0.03 for simulations that
matched the model's assumptions, versus 0.15 for simulations that included
environmental heterogeneity [[Appendix]].  For the null model, these rates were
0.31 and 0.50, respectively. Figure 4C shows, across a range of true interaction
strengths, the probability that the null model or the Markov network will
predict the wrong sign of the interaction with 95% confidence.  The null model
makes such errors nearly 8 times as often as the Markov network, even though it
only reject the null hypothesis twice as often overall [[Appendix]].  The Markov
network's errors were also more concentrated around 0, so it never misclassified
strong interactions like the null model did.

\noindent
\textbf{Discussion}

\noindent
The results presented above show that Markov networks can reliably recover
species' pairwise interactions from observational data, even for cases where
environmental heterogeneity and indirect interactions cause ecologists' typical
null modeling approaches to reliably fail. Partial covariances and generalized
linear models can both provide computationally efficient approximations, but
with somewhat lower accuracy, especially for typically-sized data sets with small
numbers of sites [@gotelli_empirical_2009]. The difference in accuracy may be
larger for real data sets than for the simulated landscapes in Figure 3,
however; linear approximations to the Markov network make larger errors when
the interaction matrix is structured [e.g. due to guilds or trophic levels;
@loh_structure_2013]. Similarly, the separate generalized linear models
for each species can severely overfit in some cases [@lee_learning_2012]. The
full Markov network should thus be preferred to the approximations when
it is computationally feasible.

Compositional data only contains enough degrees of freedom to estimate one
interaction per species pair, so none of these methods can identify the exact
nature of the pairwise interactions [e.g. which species in a
positively-associated pair is facilitating the other; cf
@schmidt_modeling_2012]. To estimate asymmetric interactions, such as
commensalism or predation, ecologists could use time series, behavioral
observations, manipulative experiments, or natural history. These other sources
of information could also be used to augment the likelihood function with an
informative prior distribution, reducing ecologists' error and uncertainty.

@@

Finally, the results presented here have important implications for
ecologists' continued use of null models for studying species interactions. Null
and neutral models can be useful for clarifying our thinking
[@harris_occupancy_2011; @xiao_strong_2015], but deviations from a particular
null model must be interpreted with care [@roughgarden_competition_1983]. Even
in small networks with three species, it may simply not be possible to implicate
specific ecological processes like competition by rejecting a general-purpose
null [@gotelli_empirical_2009], especially when the test statistic is
effectively just a correlation coefficient Figure 4A. When a non-null
backdrop is not controlled for, Type I error rates can skyrocket, the apparent
sign of the interaction can change, and null models can lead to reliably
incorrect inferences (Figure 1, Figure 4).

@@

\noindent \textbf{Figure 4: A.} The Markov network's estimated interaction
coefficients were generally very similar to the GLM estimates. **B.** The null
model's estimates typically matched the correlation coefficient, after
controlling for landscape size. **C.** For any given interaction strength, the
null model was much more likely to misclassify its sign with 95% confidence than
the Markov network was.

\noindent \textbf{References:}

### 2015-11-01 Harris_2015/manuscript-materials/draft.tex
and another thing

\documentclass[12pt,]{article}
\usepackage[compact]{titlesec}
\usepackage{fancyhdr}
\usepackage{hanging}
\pagestyle{fancy}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{setspace}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{geometry}
\usepackage{ifthen}
\geometry{verbose,letterpaper,tmargin=2.54cm, bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{4pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\subsubsection{Introduction}\label{introduction}

@@

To the extent that nontrophic species interactions (such as competition)
affect community assembly, ecologists might expect to find signatures of
these interactions in species composition data (MacArthur 1958, Diamond
1975). Despite decades of work and several major controversies, however
(Lewin 1983, Strong et al. 1984, Connor et al. 2013), existing methods
for detecting competition's effects on community structure are
unreliable (Gotelli and Ulrich 2009). In particular, species' effects on
one another can become lost in a web of indirect effects. For example,
the competitive interaction between the two shrub species in Figure 1A
is obscured by their shared tendency to occur in unshaded areas (Figure
1B). While ecologists have long known that indirect effects can
overwhelm direct ones at the landscape level (Dodson 1970, Levine 1976),
the vast majority of our methods for drawing inferences from
observational data do not control for these effects (e.g. Diamond 1975,
Strong et al. 1984, Gotelli and Ulrich 2009, Veech 2013, Pollock et al.
2014). To the extent that indirect interactions like those in Figure 1
are generally important (Dodson 1970), existing methods will not provide
much evidence regarding species interactions.

@@

\subsubsection{Methods}\label{methods}

\textbf{Markov networks.} Markov networks provide a framework for
translating back and forth between the conditional (all-else-equal)
relationships among species (Figure 1C) and the kinds of species
assemblages that these relationships produce. Here, I show how a set of
conditional relationships can be used to determine how groups of species
can co-occur. Methods for estimating conditional relationships from data
are discussed in the next section.

@@

\subsubsection{Results}\label{results}

@@

\noindent
\textbf{Twenty species.} Despite some variability across contexts
(Figure 3), the four methods that controlled for indirect effects (the
Markov network, the generalized linear models, and the two partial
correlation-based methods) outperformed those that did not. Of these,
the Markov network consistently performed the best, especially on the
smaller landscapes (which are also the most representative of typical
analyses in this field; Gotelli and Ulrich 2009). As shown in Appendix
{[}{[}X{]}{]}, the standard errors associated with the estimates in
Figure 3 are small (less than 0.01), so these differences should not be
attributed to sampling error.

Of the methods that did not control for indirect effects, Figure 3 shows
that simple correlation coefficients provided a more reliable indicator
of species' true interaction strengths than either the joint species
distribution model (BayesComm) or the null model (Pairs). The estimates
from these three approaches were tightly correlated (Figure
{[}{[}3B{]}{]}), suggesting that the null model only contains a noisy
version of the same information that could be obtained more easily and
interpretably with simple correlation coefficients.

Finally, we can evaluate the models' statistical inferences. Figure 4
shows, across a range of true interaction strengths, the probability
that the null model or the Markov network will predict the wrong sign of
the interaction with 95\% confidence. The null model makes such errors
nearly 8 times as often as the Markov network, even though it only
reject the null hypothesis twice as often overall {[}{[}Appendix{]}{]}.
The Markov network's errors were also more concentrated around 0, so it
never misclassified strong interactions like the null model did.

\subsubsection{Discussion}\label{discussion}

The results presented above show that Markov networks can reliably
recover species' pairwise interactions from observational data, even for
cases where environmental heterogeneity and indirect interactions cause
ecologists' typical null modeling approaches to reliably fail. Partial
covariances and generalized linear models can both provide
computationally efficient approximations, but with somewhat lower
accuracy. The difference in accuracy may be larger for real data sets
than for the simulated landscapes in Figure 3, however; linear
approximations to the Markov network make larger errors when the
interaction matrix is structured (e.g.~due to guilds or trophic levels;
Loh and Wainwright 2013). Similarly, the separate generalized linear
models for each species can severely overfit in some cases (Lee and
Hastie 2012). The full Markov network should thus be preferred to the
approximations when it is computationally feasible.

Compositional data only has enough degrees of freedom to estimate one
interaction per species pair (Schmidt and Murphy 2012), so none of these
methods can identify the exact nature of the pairwise interactions
{[}e.g.~which species in a positively-associated pair is facilitating
the other{]}. To estimate asymmetric interactions, such as commensalism
or predation, ecologists could use time series, behavioral observations,
manipulative experiments, or natural history. These other sources of
information could also be used to augment the likelihood function with
an informative prior distribution, reducing ecologists' eror and
uncertainty.

@@

Finally, the results presented here have important implications for
ecologists' continued use of null models for studying species
interactions. Null and neutral models can be useful for clarifying our
thinking (Harris et al. 2011, Xiao et al. 2015), but deviations from a
particular null model must be interpreted with care (Roughgarden 1983).
Even in small networks with three species, it may simply not be possible
to implicate specific ecological processes like competition by rejecting
a general-purpose null (Gotelli and Ulrich 2009), especially when the
test statistic is effectively just a correlation coefficient
{[}{[}Figure 3B{]}{]}. When a non-null backdrop is not controlled for,
Type I error rates can skyrocket, the apparent sign of the interaction
can change, and null models can lead to reliably incorrect inferences
(Figure 1, Figure 4).

@@

\noindent \textbf{Figure 4: A.} Inferences from the Markov network's
95\% confidence intervals on the ``no-environment'' simulated
landscapes, as a function of the ``true'' coefficient values.
\textbf{B.} Inferences from the \(Z\)-statistics estimated by the
``Pairs'' software for null modeling.

\textbf{References:}

>>>>

\documentclass[12pt,]{article}
\usepackage[compact]{titlesec}
\usepackage{fancyhdr}
\usepackage{hanging}
\pagestyle{fancy}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{setspace}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{geometry}
\usepackage{ifthen}
\geometry{verbose,letterpaper,tmargin=2.5cm, bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{4pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\noindent\textbf{Introduction}

@@

\noindent
To the extent that nontrophic species interactions (such as competition)
affect community assembly, ecologists might expect to find signatures of
these interactions in species composition data (MacArthur 1958, Diamond
1975). Despite decades of work and several major controversies, however
(Lewin 1983, Strong et al. 1984, Connor et al. 2013), existing methods
for detecting competition's effects on community structure are
unreliable (Gotelli and Ulrich 2009). In particular, species' effects on
one another can become lost in a web of indirect effects. For example,
the competitive interaction between the two shrub species in Figure 1A
is obscured by their shared tendency to occur in unshaded areas (Figure
1B). While ecologists have long known that indirect effects can
overwhelm direct ones at the landscape level (Dodson 1970, Levine 1976),
the vast majority of our methods for drawing inferences from
observational data do not control for these effects (e.g. Diamond 1975,
Strong et al. 1984, Gotelli and Ulrich 2009, Veech 2013, Pollock et al.
2014). To the extent that indirect interactions like those in Figure 1
are generally important (Dodson 1970), existing methods will not provide
much evidence regarding species interactions.

@@

\noindent\textbf{Methods}

\noindent
\textbf{Markov networks.} Markov networks provide a framework for
translating back and forth between the conditional (all-else-equal)
relationships among species (Figure 1C) and the kinds of species
assemblages that these relationships produce. Here, I show how a set of
conditional relationships can be used to determine how groups of species
can co-occur. Methods for estimating conditional relationships from data
are discussed in the next section.

@@

\noindent
\textbf{Results}

@@

\noindent
\textbf{Twenty species.} Despite some variability across contexts
(Figure 3), the four methods that controlled for indirect effects (the
Markov network, the generalized linear models, and the two partial
correlation-based methods) outperformed those that did not. Of these,
the Markov network consistently performed the best. As anticipated by
Lee and Hastie (2012), the Markov network produced very similar
estimates to the generalized linear models (Figure 4A), but with higher
accuracy---especially with smaller data sets. As reviewed in Gotelli and
Ulrich (2009), most analyses in this field in ecology involve fewer than
50 sites, so this is the most important regime for model performance. As
shown in Appendix {[}{[}X{]}{]}, the standard errors associated with the
estimates in Figure 3 are small (less than 0.01), so these differences
should not be attributed to sampling error.

Of the methods that did not control for indirect effects, Figure 3 shows
that simple correlation coefficients provided a more reliable indicator
of species' true interaction strengths than either the joint species
distribution model (BayesComm) or the null model (Pairs). The estimates
from these three approaches were tightly correlated, after controlling
for the size of the landscape, suggesting that the null model only
contains a noisy version of the same information that could be obtained
more easily and interpretably with simple correlation coefficients
(Figure 4B).

Finally, we can evaluate the models' statistical inferences (focusing on
the first two simulation types, for which the true interaction rates are
easiest to interpret). The Markov network's Type I error rate (defined
here as its probability of rejecting the null hypothesis for a pair of
species where \(|\beta_{ij}|<0.1\) depended on the simulation type: 0.03
for simulations that matched the model's assumptions, versus 0.15 for
simulations that included environmental heterogeneity
{[}{[}Appendix{]}{]}. For the null model, these rates were 0.31 and
0.50, respectively. Figure 4C shows, across a range of true interaction
strengths, the probability that the null model or the Markov network
will predict the wrong sign of the interaction with 95\% confidence. The
null model makes such errors nearly 8 times as often as the Markov
network, even though it only reject the null hypothesis twice as often
overall {[}{[}Appendix{]}{]}. The Markov network's errors were also more
concentrated around 0, so it never misclassified strong interactions
like the null model did.

\noindent
\textbf{Discussion}

\noindent
The results presented above show that Markov networks can reliably
recover species' pairwise interactions from observational data, even for
cases where environmental heterogeneity and indirect interactions cause
ecologists' typical null modeling approaches to reliably fail. Partial
covariances and generalized linear models can both provide
computationally efficient approximations, but with somewhat lower
accuracy, especially for typically-sized data sets with small numbers of
sites (Gotelli and Ulrich 2009). The difference in accuracy may be
larger for real data sets than for the simulated landscapes in Figure 3,
however; linear approximations to the Markov network make larger errors
when the interaction matrix is structured (e.g.~due to guilds or trophic
levels; Loh and Wainwright 2013). Similarly, the separate generalized
linear models for each species can severely overfit in some cases (Lee
and Hastie 2012). The full Markov network should thus be preferred to
the approximations when it is computationally feasible.

Compositional data only contains enough degrees of freedom to estimate
one interaction per species pair, so none of these methods can identify
the exact nature of the pairwise interactions (e.g.~which species in a
positively-associated pair is facilitating the other; cf Schmidt and
Murphy 2012). To estimate asymmetric interactions, such as commensalism
or predation, ecologists could use time series, behavioral observations,
manipulative experiments, or natural history. These other sources of
information could also be used to augment the likelihood function with
an informative prior distribution, reducing ecologists' error and
uncertainty.

@@

Finally, the results presented here have important implications for
ecologists' continued use of null models for studying species
interactions. Null and neutral models can be useful for clarifying our
thinking (Harris et al. 2011, Xiao et al. 2015), but deviations from a
particular null model must be interpreted with care (Roughgarden 1983).
Even in small networks with three species, it may simply not be possible
to implicate specific ecological processes like competition by rejecting
a general-purpose null (Gotelli and Ulrich 2009), especially when the
test statistic is effectively just a correlation coefficient Figure 4A.
When a non-null backdrop is not controlled for, Type I error rates can
skyrocket, the apparent sign of the interaction can change, and null
models can lead to reliably incorrect inferences (Figure 1, Figure 4).

@@

\noindent \textbf{Figure 4: A.} The Markov network's estimated
interaction coefficients were generally very similar to the GLM
estimates. \textbf{B.} The null model's estimates typically matched the
correlation coefficient, after controlling for landscape size.
\textbf{C.} For any given interaction strength, the null model was much
more likely to misclassify its sign with 95\% confidence than the Markov
network was.

\noindent \textbf{References:}
